Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences

### 2020-09-30 manuscript/manuscript.Rmd
Final minor R3 edits

---
title: "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"
short-title: Dynamics of International Giving
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State restrictions on non-governmental organizations (NGOs) have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private philanthropy. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and main funding sources. We discuss the implications of our results on how nonprofits can use different framing appeals to increase fundraising at a time when traditional public donor funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,012"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
---

@@

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. We use a survey experiment fielded in the U.S. to assess how different structural and organizational-level heuristics affect donor preferences. We find that learning about repressive NGO environments increases generosity and that already-likely donors are willing to donate substantially more to legally restricted nonprofits.

@@

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard [@ChongDruckman:2007]. However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how this frame interacts with organizational attributes to change donor preferences. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though we focus on private funding in this article, it is essential to differentiate between the two sources. Public aid, or more traditional donor aid, channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds come from foundations, corporations, and individuals. We are specifically interested in individual-level determinants of private international philanthropy.

Private giving to nonprofits working in international affairs has increased consistently over the past decade. Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international nonprofits, they accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors thus play an incredibly important—and underexplored—role in funding international NGOs.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different frames donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://bit.ly/osf-prereg-anon](https://bit.ly/osf-prereg-anon). The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, CONSORT diagram, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

>>>>

---
title: "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"
short-title: Dynamics of International Giving
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "September 29, 2020"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State restrictions on non-governmental organizations (NGOs) have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private philanthropy. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and main funding sources. We discuss the implications of our results on how nonprofits can use different framing appeals to increase fundraising at a time when traditional public donor funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,012"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
---

@@

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard [@ChongDruckman:2007], as NGOs' appeals for donations are framed by structural and organizational characteristics. We use a survey experiment fielded in the U.S. to assess how different structural and organizational-level heuristics affect donor preferences. We find that learning about repressive NGO environments increases generosity and that already-likely donors are willing to donate substantially more to legally restricted nonprofits.

@@

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. In the era of closing civic space, NGOs may need to reframe and tailor their fundraising strategies to individual donors. However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how this frame interacts with organizational attributes to change donor preferences. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs (INGOs) in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though we focus on private funding in this article, it is essential to differentiate between the two sources. Public aid, or more traditional donor aid, channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds come from foundations, corporations, and individuals. We are specifically interested in individual-level determinants of private international philanthropy.

Nonprofits working internationally typically rely heavily on official aid from government sources. In 2018, for instance, humanitarian INGOs received 81% of their total funding from governments and EU institutions [@Development-Initiatives:2020]. However, private giving to nonprofits working in international affairs has increased consistently over the past decade. Donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international nonprofits, they accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Moreover, while government, foundation, and corporate funding is often earmarked for specific purposes, private donations are typically unrestricted funds.[^reviewer-thanks] Individual donors thus play an incredibly important—and underexplored—role in funding international NGOs.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different frames donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://doi.org/10.17605/OSF.IO/DX973>. The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, CONSORT diagram, and Bayesian priors. The analysis is fully reproducible using code and data available at <https://github.com/andrewheiss/ngo-crackdowns-philanthropy> and <https://dx.doi.org/10.17605/OSF.IO/FG53W>.

@@

[^reviewer-thanks]: We thank an anonymous reviewer for this insight.
